A Historical Analysis of the Evolution of Active Travel Behaviour in Canada

### 2024-09-22 papers/article/active-travel-behaviour-evolution-in-Canada/active-travel-behaviour-evolution-in-Canada.Rmd
Adding the rmd with the methodology and already-processed data

The merit of this function is due to its ability to assign decreasing influences to more remote opportunities, giving a more accurate estimate for shorter journeys [@iacono2010; @kanafani1983transportation; @fotheringham1989spatial]. However, in addition to determine the form of the impedance function, scholars also need to specify the variable used to measure impedance, which can be either time, distance, monetary cost, a combination these last variables or even a generalized cost concept.

The choice between time and distance as the impedance variable has been found to be acceptable based on previous research [@iacono2010; @hull2012accessibility; @sun2012measuring; @lowry2012using; @vasconcelos2012evaluation ]. However, when it comes to non-motorized travel modes, extracting accurate travel times from existing network models can be challenging, which limits the options and makes distance a more practical choice [@handy1997measuring; @iacono2010; @yang2012walking; @arranz2019measuring]. Furthermore, researchers specializing in active modes of transportation have faced challenges stemming from a need for more objective data concerning walking and cycling behavior. Estimating specialized impedance functions specific to non-motorized modes requires appropriate travel survey data that can capture pedestrian and cycle behavior. Often, researchers have resorted to relying on retrospective questionnaires, which assess subjective aspects such as the frequency and duration of walking and cycling activities. Notably, regional household travel surveys, including trips made by non-motorized modes, have been employed for this purpose [@iacono2010; @millward2013active].

In contrast to these localized surveys, some datasets provide a nationwide perspective, encompassing travel for various trip purposes and offering insights into details like travel episode origins, destinations, and time-based lengths. This comprehensive approach furnishes a more holistic understanding of active transportation behavior. Nevertheless, only some studies, such as Yang et al. (2012), have examined active travel behavior nationally. Addressing the previously identified gaps and precisely the challenge of using uniform decay curves, such as the negative exponential function for both cycling and walking in accessibility assessments, this study endeavors to identify the most appropriate distance decay function for various destinations and time periods.  Our research delves into the complexities of actual travel behavior, emphasizing active transportation modes. We utilize historical data from the General Social Survey (GSS) from 1986 to 2015 in Canada to calculate the impedance function for cycling and walking trips. By doing so, this study conducts a comparative analysis of travel behaviors associated with these two modes.

Moreover, we recognize that non-work travel encompasses a spectrum of trip intentions and diverse traveler behaviors, making the impedance function an essential analytical tool for examining non-work accessibility. Grengs (2015) emphasizes the importance of crafting distinct functions for each travel purpose, a principle that guides our analysis [@grengs2015nonwork]. Our investigation will encompass a variety of trip intents, ranging from commutes to homes, workplaces, or educational institutions to social visits, outdoor activities, business trips, shopping, cultural outings to libraries, museums, or theaters, dining out, and engaging in religious practices. By drawing on a nationally representative sample of Canadian residents, our research intends to bridge the empirical data gap concerning the frequency and duration of typical pedestrian and cycling trips for different purposes. In doing so, it aspires to provide a nuanced understanding of active travel behavior. Ultimately, this comprehensive analysis seeks to contribute meaningfully to the ongoing conversation on active transportation, shedding light on its role in shaping travel behavior and accessibility.

>>>>

The merit of this function is due to its ability to assign decreasing influences to more remote opportunities, giving a more accurate estimate for shorter trips [@iacono2010; @kanafani1983transportation; @fotheringham1989spatial]. However, in addition to determine the form of the impedance function, scholars also need to specify the variable used to measure impedance, which can be either time, distance, monetary cost, a combination these last variables or even a generalized cost concept. Among these options, the choice between time and distance as the impedance has been found to be most used based on previous studies [@iacono2010; @hull2012accessibility; @sun2012measuring; @lowry2012using; @vasconcelos2012evaluation], with distance being more adopted in non-motorized applications since extracting accurate travel times from existing network models is challenging [@handy1997measuring; @iacono2010; @yang2012walking; @arranz2019measuring]. Additionally, estimate impedance function to active transportation modes requires appropriate travel survey data that is able to capture pedestrian and cycle behaviour, resulting in researchers recurring to retrospective questionnaires to assess subjective aspects such as the frequency and duration of walking and cycling activities. Notably, regional household travel surveys that include trips made by non-motorized modes have been employed for this purpose [@iacono2010; @millward2013active]. In opposition to these specific surveys, some data sets provides a nationwide perspective, including travel for different purposes and detailing the trip with valuable information, named episodes, regarding the origins, destinations, and time-based lengths. Besides this type of data can provides a deeper comprehension about the active transportation behaviour,  such as Yang et al. (2012), only few studies have examined travel behaviour nationally. 
It is aiming to address the mentioned challenges that this study focuses on identifying the most appropriate impedance functions for active transportation modes for various destinations and time periods in Canada. To do achieve our objective, we analyze historical data from the General Social Survey (GSS) from 1986 to 2015 in Canada to calculate the impedance function for cycling and walking trips. Additionally, this study perform a comparative analysis of travel behaviors associated with these two modes.

We recognize that non-work travel encompasses a range of trip purposes and diverse traveler behaviors, which makes impedance functions essential analytical tools for studying non-work accessibility. Grengs (2015) emphasizes the importance of elaborating distinct functions for each travel purpose, a principle that guides this analysis [@grengs2015nonwork]. Our investigation covers a variety of trip goals, ranging from commutes to homes, workplaces, or educational institutions to social visits, outdoor activities, business trips, shopping, cultural outings to libraries, museums, or theaters, dining out, and engaging in religious practices. 
Our research aims to enhance the current knowledge about active travel behaviour and provide empirical data about frequency and duration of typical pedestrian and cycling trips for different purposes, by applying the methodology on a nationally representative samples of Canadian residents. Lastly, this comprehensive analysis seeks to contribute to the ongoing conversation on active transportation, highlighting its role in influencing transportation plans to a more sustainable alternative.
